---
title: "idealized Design Matrix"
author: "Caryn Johansen"
date: "10/27/2017"
output:
  pdf_document: default
  html_document: default
---

```{r}
library(Matrix)
library(tidyverse)
library(lme4)
```

# Full design Matrix

```{r}
dm2 <- read.table("designM_ideal.csv", header=T, se=",")
image(Matrix(as.matrix(dm2[,-c(1:4)])))
dim(dm2)
```
```{r}
table(dm2$Allele)
table(dm2$Cross)
```

random tpm data
```{r}
y <- rnorm(1e4, mean = 0.679, sd = 0.8)
y <- subset(y, y>-0.3)
random_tpm_log <- sample(y, nrow(dm2))
random_tpm <- 10**random_tpm_log
dm2$tpm <- random_tpm
```


Full Model

```{r}
dm2_full_fit1 <- lmer(tpm ~ (1|PlantID) + cis_s + cis_m + trans_m + trans_s +
                       trans_ms + 
                       cis_m_trans_m + cis_m_trans_s +
                       cis_s_trans_m + cis_s_trans_s +
                       cis_m_trans_ms +
                       cis_s_trans_ms, data=dm2)
summary(dm2_full_fit1)
full_mm <- model.matrix(tpm ~ cis_s + cis_m + trans_m + trans_s +
                       trans_ms + trans_mb + trans_sb +
                       cis_m_trans_m + cis_m_trans_s +
                       cis_s_trans_m + cis_s_trans_s +
                       cis_m_trans_ms +
                       cis_s_trans_ms, data=dm2)
lin_combos <- caret::findLinearCombos(full_mm)
lin_combos
colnames(full_mm)
```

## Simulating effect

```{r}
X <- as.matrix(dm2[,-c(1:4,29)])
```


### cis m

```{r message=FALSE, warning=FALSE}
p_values_list <- list()

for ( i in 1:ncol(X)) {

  b <- rep(0,ncol(X))
  b[i] <- 1
  
  p_values <- c()
  y <- X %*% b + rnorm(nrow(X), mean=0, sd=1e-10)
  tryCatch({
    fit1 <- lmer(y ~ (1|PlantID) + cis_s + cis_m + trans_m + trans_s + 
                   trans_ms + 
                   cis_m_trans_m + cis_m_trans_s +
                   cis_s_trans_m + cis_s_trans_s +
                   cis_m_trans_ms +
                   cis_s_trans_ms, data=dm2)
    }, error=function(c) NA)
    tryCatch({
       fit2 <- lmer(y ~ (1|PlantID) + cis_s + trans_m + trans_s +
                       trans_ms + 
                       cis_s_trans_m + cis_s_trans_s +
                       cis_s_trans_ms, data=dm2)
    }, error=function(e) NA)
    tryCatch({
      p <- anova(fit1, fit2)$`Pr(>Chisq)`[2]
    }, error=function(e) p <- NA)
    p_values <- c(p_values, p)
  p_values <- data.frame(variable=rep(colnames(X)[i],length(p_values)), p_values=p_values)
  p_values_list[[i]] <- p_values
}

p_cis_m <- do.call("rbind", p_values_list)
p_cis_m$sigTF <- p_cis_m$p_values < 1e-10
ggplot(p_cis_m) + geom_point(aes(x=as.factor(variable), y=p_values, color=sigTF)) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### cis s

```{r message=FALSE, warning=FALSE}
p_values_list <- list()
#p_values <- c()
for ( i in 1:ncol(X)) {

  b <- rep(0,ncol(X))
  b[i] <- 1
  
  y <- X %*% b + rnorm(nrow(X), mean=0, sd=1e-10)
  tryCatch({
    fit1 <- lmer(y ~ (1|PlantID) + cis_s + cis_m + trans_m + trans_s +
                   trans_ms + 
                   cis_m_trans_m + cis_m_trans_s +
                   cis_s_trans_m + cis_s_trans_s +
                   cis_m_trans_ms +
                   cis_s_trans_ms, data=dm2)
    }, error=function(c) NA)
  tryCatch({
    fit2 <- lmer(y ~ (1|PlantID) + cis_m + trans_m + trans_s +
                     trans_ms + 
                     cis_m_trans_m + cis_m_trans_s +
                     cis_m_trans_ms, data=dm2)
  }, error=function(e) NA)
  tryCatch({
      p <- anova(fit1, fit2)$`Pr(>Chisq)`[2]
  }, error=function(e) p <- NA)
  p_values <- data.frame(variable=rep(colnames(X)[i],length(p)), p_values=p)
  p_values_list[[i]] <- p_values
}

p_cis_s <- do.call("rbind", p_values_list)
p_cis_s$sigTF <- p_cis_s$p_values < 1e-10
ggplot(p_cis_s) + geom_point(aes(x=as.factor(variable), y=p_values, color=sigTF)) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


### trans ms

remove interacting values

```{r message=FALSE, warning=FALSE}
p_values_list <- list()

for ( i in 1:ncol(X)) {

  b <- rep(0,ncol(X))
  b[i] <- 1
  
  y <- X %*% b + rnorm(nrow(X), mean=0, sd=1e-10)
  tryCatch({
    fit1 <- lmer(y ~ (1|PlantID) + cis_s + cis_m + trans_m + trans_s +
                       trans_ms + 
                       cis_m_trans_m + cis_m_trans_s +
                       cis_s_trans_m + cis_s_trans_s +
                       cis_m_trans_ms +
                       cis_s_trans_ms, data=dm2)
  }, error=function(c) NA)
  tryCatch({
    fit2 <- lmer(y ~ (1|PlantID) + cis_s + cis_m + trans_m + trans_s +
                       cis_m_trans_m + #cis_m_trans_s +
                       cis_s_trans_s, data=dm2)
  }, error=function(e) NA)
  tryCatch({
    p <- anova(fit1, fit2)$`Pr(>Chisq)`[2]
  }, error=function(e) p <- NA)
  
  p_values <- data.frame(variable=rep(colnames(X)[i],length(p)), p_values=p)
  p_values_list[[i]] <- p_values
}

p_trans <- do.call("rbind", p_values_list)
p_trans$sigTF <- p_trans$p_values < 1e-10
ggplot(p_trans) + geom_point(aes(x=as.factor(variable), y=p_values, color=sigTF)) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Without B73

Remove the inbred B73 lines, which are the first three plants (hence the first 6 rows)

```{r}
dm3 <- dm2[-c(1:6),]
image(Matrix(as.matrix(X)))
X2 <- X[-c(1:6),]
image(Matrix(as.matrix(X2)))
```

Run the simulation again:

```{r message=FALSE, warning=FALSE}
p_values_list <- list()

for ( i in 1:ncol(X2)) {

  b <- rep(0,ncol(X2))
  b[i] <- 1
  
  p_values <- c()
  y <- X2 %*% b + rnorm(nrow(X2), mean=0, sd=1e-10)
  tryCatch({
    fit1 <- lmer(y ~ (1|PlantID) + cis_s + cis_m + trans_m + trans_s +
                       trans_ms + 
                       cis_m_trans_m + cis_m_trans_s +
                       cis_s_trans_m + cis_s_trans_s +
                       cis_m_trans_ms +
                       cis_s_trans_ms, data=dm3)
  }, error=function(c) NA)
  tryCatch({
       fit2 <- lmer(y ~ (1|PlantID) + cis_s + cis_m + trans_m + trans_s +
                       cis_m_trans_m + #cis_m_trans_s +
                       cis_s_trans_s, data=dm3)
  }, error=function(e) NA)
  tryCatch({
    p <- anova(fit1, fit2)$`Pr(>Chisq)`[2]
  }, error=function(e) p <- NA)
  
  p_values <- data.frame(variable=rep(colnames(X2)[i],length(p)), p_values=p)
  p_values_list[[i]] <- p_values
}

p_trans_noB <- do.call("rbind", p_values_list)
p_trans_noB$sigTF <- p_trans_noB$p_values < 1e-10
ggplot(p_trans_noB) + geom_point(aes(x=as.factor(variable), y=p_values, color=sigTF)) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### No landrace crosses

Remove MEXxMEX and SAMxSAM, and just look at B73, and the crosses between landraces

```{r}
dm4 <- dm2 %>% filter(Cross != "MEXxMEX", Cross != "SAMxSAM")
X4 <- as.matrix(dm4[,-c(1:4,29)])
```

```{r message=FALSE, warning=FALSE}
p_values_list <- list()

for ( i in 1:ncol(X4)) {

  b <- rep(0,ncol(X4))
  b[i] <- 1
  
  y <- X4 %*% b + rnorm(nrow(X4), mean=0, sd=1e-10)
  tryCatch({
      fit1 <- lmer(y ~ (1|PlantID) + cis_s + cis_m + trans_m + trans_s +
                       trans_ms + 
                       cis_m_trans_m + cis_m_trans_s +
                       cis_s_trans_m + cis_s_trans_s +
                       cis_m_trans_ms +
                       cis_s_trans_ms, data=dm4)
  }, error=function(c) NA)
  tryCatch({
       fit2 <- lmer(y ~ (1|PlantID) + cis_s + cis_m + trans_m + trans_s +
                       cis_m_trans_m + #cis_m_trans_s +
                       cis_s_trans_s, data=dm4)
  }, error=function(e) NA)
  tryCatch({
      p <- anova(fit1, fit2)$`Pr(>Chisq)`[2]
  }, error=function(e) p <- NA)

  p_values <- data.frame(variable=rep(colnames(X4)[i],length(p)), p_values=p)
  p_values_list[[i]] <- p_values
}

p_trans_noMMSS <- do.call("rbind", p_values_list)
p_trans_noMMSS$sigTF <- p_trans_noMMSS$p_values < 1e-10
ggplot(p_trans_noMMSS) + geom_point(aes(x=as.factor(variable), y=p_values, color=sigTF)) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```




### Difference between cis_b and cis_m/cis_s

Attempting to distinguish between cis_s/cis_m and cis_b

```{r}
X <- as.matrix(dm2[,-c(1:4,29)]) #making sure I don't use an altered X that would mess up my interpretation
```

Simulating cis_b

```{r message=FALSE, warning=FALSE}
p_values_list <- list()


for (i in 1:ncol(X)) {
  b <- rep(0,ncol(X))
  b[i] <- 1

  y <- X %*% b + rnorm(nrow(X), mean=0, sd=1e-10)
  tryCatch({
      fit1 <- lmer(y ~ (1|Cross2) + cis_s + cis_m + trans_m + trans_s +
                       trans_ms +
                       cis_m_trans_m + cis_m_trans_s +
                       cis_s_trans_m + cis_s_trans_s +
                       cis_m_trans_ms +
                       cis_s_trans_ms, data=dm2)
  }, error=function(c) NA)
  tryCatch({
      fit2 <- lmer(y ~ (1|Cross2) + cis_s + trans_m + trans_s +
                       trans_ms +
                       cis_s_trans_m + cis_s_trans_s +
                       cis_s_trans_ms, data=dm2)
  }, error=function(e) NA)
  tryCatch({
      p <- anova(fit1, fit2)$`Pr(>Chisq)`[2]
  }, error=function(e) p <- NA)

  p_values <- data.frame(variable=rep(colnames(X)[i],length(p)), p_values=p)
  p_values_list[[i]] <- p_values
} 

p_cis_m2 <- do.call("rbind", p_values_list)
p_cis_m2$sigTF <- p_cis_m2$p_values < 1e-10
ggplot(p_cis_m2) + geom_point(aes(x=as.factor(variable), y=p_values, color=sigTF)) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


##### Without landraces

```{r message=FALSE, warning=FALSE}
p_values_list <- list()


for ( i in 1:ncol(X4)) {

  b <- rep(0,ncol(X4))
  b[i] <- 1

  y <- X4 %*% b + rnorm(nrow(X4), mean=0, sd=1e-10)
  tryCatch({
      fit1 <- lmer(y ~ (1|PlantID) + cis_s + cis_m + trans_m + trans_s +
                       trans_ms +
                       cis_m_trans_m + cis_m_trans_s +
                       cis_s_trans_m + cis_s_trans_s +
                       cis_m_trans_ms +
                       cis_s_trans_ms, data=dm4)
  }, error=function(c) NA)
  tryCatch({
       fit2 <- lmer(y ~ (1|PlantID) + cis_s + trans_m + trans_s +
                       trans_ms +
                       cis_s_trans_m + cis_s_trans_s +
                       cis_s_trans_ms, data=dm4)
  }, error=function(e) NA)
  tryCatch({
      p <- anova(fit1, fit2)$`Pr(>Chisq)`[2]
  }, error=function(e) p <- NA)

  p_values <- data.frame(variable=rep(colnames(X4)[i],length(p)), p_values=p)
  p_values_list[[i]] <- p_values
}

p_cis_m3 <- do.call("rbind", p_values_list)
p_cis_m3$sigTF <- p_cis_m3$p_values < 1e-10
ggplot(p_cis_m3) + geom_point(aes(x=as.factor(variable), y=p_values, color=sigTF)) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


# Condense visual of models

Combine sigTF columns into one data frame. Geom_tile

```{r}
p_trans_model <- data.frame(model=rep("p_trans",nrow(p_trans)), parameter=p_trans$variable ,sigTF = p_trans$sigTF)
p_cis_m_model <- data.frame(model=rep("p_cis_m",nrow(p_cis_m)), parameter=p_cis_m$variable, sigTF = p_cis_m$sigTF)

models <- list()
models[[1]] <- p_trans_model
models[[2]] <- p_cis_m_model
mods <- do.call("rbind", models)
ggplot(mods, aes(x=parameter, y=model)) + geom_tile(aes(fill=sigTF)) + theme(axis.text.x = element_text(angle=45, hjust = 1))
```


# With Current DM

Restructure DM on Mex allele


```{r}
d <- data.frame(cis=rep(c("B","S","M"),2))
model.matrix(~cis,d)
```

```{r}
d$T1 <- rep(c("BS","BM","MS"), each=2)

d$TM <- c(0,1)[grepl('M',d$T1)+1]
d$TS <- c(0,1)[grepl('S', d$T1)+1]
d$TB <- c(0,1)[grepl('B', d$T1)+1]

dm1 <- as.data.frame(model.matrix(~cis*TM*TS,d))
dm1 <- as.data.frame(rbind(dm1, dm1, dm1))
dm1$allele <- rep(d$cis, 3)
dm1$cross <- rep(d$T1,3)
dm1$PlantID <- rep(1:9, each=2)
dm1
```

```{r}
image(Matrix(as.matrix(dm1[,-c(13:15)])))
```


Make relative to Mexico

```{r}
dMex <- d
dMex$cis <- relevel(dMex$cis, 2) # make Mex the intercept
model.matrix(~cis*TB*TS, dMex)
```

X is the dm without the allele, cross and plant info

```{r}
X1 <- as.matrix(dm1[,-c(13:15)])
```


#### Full defined 

Where the model to use as the comparison will be what I've been calling the full model:

$$y \sim (1|PlantID) + cis_S + cis_M + trans_M + trans_S + trans_{MS} + cis_Mtrans_M + cis_Mtrans_S + cis_Strans_S + cis_Strans_M + cis_Mtrans_{MS} + cis_Strans_{MS}$$

### Alternative models

Interaction terms between M and S genomes

```{r message=FALSE, warning=FALSE}
p_values_list <- list()

for (i in 1:ncol(X1)) {
  # simulating the effect for that model component
  b <- rep(0, ncol(X1))
  b[i] <- 1
  y <- X1 %*% b + rnorm(nrow(X1), mean=0, sd=1e-10)
  full <- lmer(y ~ (1|PlantID) + cisS + cisM + TM + TS + TM:TS + cisM:TM + cisM:TS + cisS:TS + cisS:TM + cisM:TM:TS + cisS:TM:TS, data=dm1 )
  alt <- lmer(y ~ (1|PlantID) + cisS + cisM + TM + TS + cisM:TM + cisS:TS, data=dm1)
  p <- anova(full, alt)$`Pr(>Chisq)`[2]
  p_value <- data.frame(variable=colnames(X1)[i], p_value=p)
  p_values_list[[i]] <- p_value
}
p_interact <- do.call("rbind", p_values_list)
p_interact$sigTF <- p_interact$p_value < 1e-10
p_interact$model <- rep("interaction", nrow(p_interact))
#ggplot(p_interact) + geom_point(aes(x=as.factor(variable), y=p_value, color=sigTF)) +
#  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Cis S

```{r message=FALSE, warning=FALSE}
p_values_list <- list()

for (i in 1:ncol(X1)) {
  # simulating the effect for that model component
  b <- rep(0, ncol(X1))
  b[i] <- 1
  y <- X1 %*% b + rnorm(nrow(X1), mean=0, sd=1e-10)
  full <- lmer(y ~ (1|PlantID) + cisS + cisM + TM + TS + TM:TS + cisM:TM + cisM:TS + cisS:TS + cisS:TM + cisM:TM:TS + cisS:TM:TS, data=dm1 )
  alt <- lmer(y ~ (1|PlantID)  + cisM + TM + TS + TM:TS + cisM:TM + cisM:TS + cisM:TM:TS, data=dm1)
  p <- anova(full, alt)$`Pr(>Chisq)`[2]
  p_value <- data.frame(variable=colnames(X1)[i], p_value=p)
  p_values_list[[i]] <- p_value
}
p_cisS <- do.call("rbind", p_values_list)
p_cisS$sigTF <- p_cisS$p_value < 1e-10
p_cisS$model <- rep("cisS", nrow(p_cisS))
#ggplot(p_cisS) + geom_point(aes(x=as.factor(variable), y=p_value, color=sigTF)) + 
#  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Cis M

```{r message=FALSE, warning=FALSE}
p_values_list <- list()

for (i in 1:ncol(X1)) {
  # simulating the effect for that model component
  b <- rep(0, ncol(X1))
  b[i] <- 1
  y <- X1 %*% b + rnorm(nrow(X1), mean=0, sd=1e-10)
  full <- lmer(y ~ (1|PlantID) + cisS + cisM + TM + TS + TM:TS + cisM:TM + cisM:TS + cisS:TS + cisS:TM + cisM:TM:TS + cisS:TM:TS, data=dm1 )
  alt <- lmer(y ~ (1|PlantID)  + cisS + TM + TS + TM:TS + cisS:TM + cisS:TS + cisS:TM:TS, data=dm1)
  p <- anova(full, alt)$`Pr(>Chisq)`[2]
  p_value <- data.frame(variable=colnames(X1)[i], p_value=p)
  p_values_list[[i]] <- p_value
}
p_cisM <- do.call("rbind", p_values_list)
p_cisM$sigTF <- p_cisM$p_value < 1e-10
p_cisM$model <- rep("cisM", nrow(p_cisM))
#ggplot(p_interact) + geom_point(aes(x=as.factor(variable), y=p_value, color=sigTF)) +
#theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Visualize models and p-values

```{r}
modComparisons <- rbind(p_cisM, p_cisS, p_interact)
modComparisons %>% ggplot(aes(x=variable, y=model, fill=sigTF)) + geom_tile() + 
  theme(axis.text.x = element_text(angle=45, hjust=1))
```

# With B73

```{r}
d <- data.frame(cis=c("B","B",rep(c("B","S","M"), 2)))
d$T1 <- rep(c("BB","BS","BM","MS"), each=2)

d$TB <- c(0,1)[grepl('B', d$T1) + 1]
d$TM <- c(0,1)[grepl('M', d$T1) + 1]
d$TS <- c(0,1)[grepl('S', d$T1) + 1]
d
```

```{r}
dm <- model.matrix(~cis*TM*TS,d)


dm <- as.data.frame(rbind(dm, dm, dm))
dm$allele <- rep(d$cis, 3)
dm$cross <- rep(d$T1,3)
dm$PlantID <- rep(1:12, each=2)

image(Matrix(as.matrix(dm[,-c(13:15)])))
```

#### Full defined 

Where the model to use as the comparison will be what I've been calling the full model:

$$y \sim (1|PlantID) + cis_S + cis_M + trans_M + trans_S + trans_{MS} + cis_Mtrans_M + cis_Mtrans_S + cis_Strans_S + cis_Strans_M + cis_Mtrans_{MS} + cis_Strans_{MS}$$

### Alternative models

```{r}
X <- as.matrix(dm[,-c(13:15)])
```


Interaction terms between M and S genomes

```{r message=FALSE, warning=FALSE}
p_values_list <- list()

for (i in 1:ncol(X)) {
  # simulating the effect for that model component
  b <- rep(0, ncol(X))
  b[i] <- 1
  y <- X %*% b + rnorm(nrow(X), mean=0, sd=1e-10)
  full <- lmer(y ~ (1|PlantID) + cisS + cisM + TM + TS + TM:TS + cisM:TM + cisM:TS + cisS:TS + cisS:TM + cisM:TM:TS + cisS:TM:TS, data=dm )
  alt <- lmer(y ~ (1|PlantID) + cisS + cisM + TM + TS + cisM:TM + cisS:TS, data=dm)
  p <- anova(full, alt)$`Pr(>Chisq)`[2]
  p_value <- data.frame(variable=colnames(X)[i], p_value=p)
  p_values_list[[i]] <- p_value
}
p_interact <- do.call("rbind", p_values_list)
p_interact$sigTF <- p_interact$p_value < 1e-10
p_interact$model <- rep("interaction", nrow(p_interact))
#ggplot(p_interact) + geom_point(aes(x=as.factor(variable), y=p_value, color=sigTF)) +
#  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Cis S

```{r message=FALSE, warning=FALSE}
p_values_list <- list()

for (i in 1:ncol(X)) {
  # simulating the effect for that model component
  b <- rep(0, ncol(X))
  b[i] <- 1
  y <- X %*% b + rnorm(nrow(X), mean=0, sd=1e-10)
  full <- lmer(y ~ (1|PlantID) + cisS + cisM + TM + TS + TM:TS + cisM:TM + cisM:TS + cisS:TS + cisS:TM + cisM:TM:TS + cisS:TM:TS, data=dm )
  alt <- lmer(y ~ (1|PlantID)  + cisM + TM + TS + TM:TS + cisM:TM + cisM:TS + cisM:TM:TS, data=dm)
  p <- anova(full, alt)$`Pr(>Chisq)`[2]
  p_value <- data.frame(variable=colnames(X)[i], p_value=p)
  p_values_list[[i]] <- p_value
}
p_cisS <- do.call("rbind", p_values_list)
p_cisS$sigTF <- p_cisS$p_value < 1e-10
p_cisS$model <- rep("cisS", nrow(p_cisS))
#ggplot(p_cisS) + geom_point(aes(x=as.factor(variable), y=p_value, color=sigTF)) + 
#  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Cis M

```{r message=FALSE, warning=FALSE}
p_values_list <- list()

for (i in 1:ncol(X)) {
  # simulating the effect for that model component
  b <- rep(0, ncol(X))
  b[i] <- 1
  y <- X %*% b + rnorm(nrow(X), mean=0, sd=1e-10)
  full <- lmer(y ~ (1|PlantID) + cisS + cisM + TM + TS + TM:TS + cisM:TM + cisM:TS + cisS:TS + cisS:TM + cisM:TM:TS + cisS:TM:TS, data=dm )
  alt <- lmer(y ~ (1|PlantID)  + cisS + TM + TS + TM:TS + cisS:TM + cisS:TS + cisS:TM:TS, data=dm)
  p <- anova(full, alt)$`Pr(>Chisq)`[2]
  p_value <- data.frame(variable=colnames(X)[i], p_value=p)
  p_values_list[[i]] <- p_value
}
p_cisM <- do.call("rbind", p_values_list)
p_cisM$sigTF <- p_cisM$p_value < 1e-10
p_cisM$model <- rep("cisM", nrow(p_cisM))
#ggplot(p_interact) + geom_point(aes(x=as.factor(variable), y=p_value, color=sigTF)) +
#theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Visualize models and p-values

```{r}
modComparisons2 <- rbind(p_cisM, p_cisS, p_interact)
modComparisons2 %>% ggplot(aes(x=variable, y=model, fill=sigTF)) + geom_tile() + 
  theme(axis.text.x = element_text(angle=45, hjust=1))
```


```{r message=FALSE, warning=FALSE}
p_values_list <- list()

for (i in 1:ncol(X)) {
  # simulating the effect for that model component
  b <- rep(0, ncol(X))
  b[i] <- 1
  y <- X %*% b + rnorm(nrow(X), mean=0, sd=1e-10)
  full <- lmer(y ~ (1|PlantID) + cisS + cisM + TM + TS + TM:TS + cisM:TM + cisM:TS + cisS:TS + cisS:TM + cisM:TM:TS + cisS:TM:TS, data=dm )
  alt <- lmer(y ~ (1|PlantID)  + cisS + cisM + TM + TS + TM:TS + cisM:TM + cisM:TS + cisS:TS + cisS:TM + cisM:TM:TS + cisS:TM:TS, data=dm)
  p <- anova(full, alt)$`Pr(>Chisq)`[2]
  p_value <- data.frame(variable=colnames(X)[i], p_value=p)
  p_values_list[[i]] <- p_value
}
p_cisM <- do.call("rbind", p_values_list)
p_cisM$sigTF <- p_cisM$p_value < 1e-10
p_cisM$model <- rep("cisM", nrow(p_cisM))
#ggplot(p_interact) + geom_point(aes(x=as.factor(variable), y=p_value, color=sigTF)) +
#theme(axis.text.x = element_text(angle = 45, hjust = 1))
```